---
title: "Exploratory Data Analysis"
author: "Pascal Jerney"
date: "2021-02-27"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, echo=FALSE}

if (!require("pacman", quietly = TRUE)) {
  install.packages("pacman")
}
pacman::p_load(
  "tidyverse",
  "conflicted",
  "fs",
  "glue",
  "ggridges",
  "data.table"
)

conflict_prefer("select", "dplyr", quiet = TRUE)
conflict_prefer("filter", "dplyr", quiet = TRUE)

one_over_mac <-
  read_rds(file = path_wd("output", "one_over_mac", ext = "rds"))

one_over_mac_filtered <-
  read_rds(file = path_wd("output", "one_over_mac_filtered", ext = "rds"))

`%!in%` <- Negate(`%in%`)

```

## Introduction

## To do

* Add density adjustments and/or histogram to univariate plots
* Use stat_density for density plots
* Rename univariate density plots
* Add univariate age plots
* Add univariate natural logarithm of ce_mac plot

## Data structure

```{r data_structure}

one_over_mac[1:20, .(pid, age, time, ce_mac, peak_alpha)] %>%
  knitr::kable(col.names = c("Patient ID", "Age (years)", "Time (sec)",
                             "ceMAC", "Peak alpha frequency"),
               caption = "Raw data")

melt(one_over_mac[,
                  lapply(.SD,
                         function(x)
                           sum(is.na(x))),
                  .SDcols = c("pid", "age", "time", "ce_mac", "peak_alpha")],
     measure.vars = c("pid", "age", "time", "ce_mac", "peak_alpha"),
     value.name = "count") %>%
  knitr::kable(col.names = c("Variable", "Count"),
               caption = "Missing values")

quartiles <-
  one_over_mac[, .N, by = pid
               ][, quantile(N, seq(0.25, .75, by = 0.25))]

one_over_mac[, .N, by = pid
             ][, .(x = density(N)$x, y = density(N)$y)
               ][, .(x, y,
                     quartile = factor(findInterval(x, quartiles) + 1))] %>%
  ggplot(aes(x = x, y = y)) +
  geom_line() +
  geom_ribbon(aes(ymin = 0, ymax = y, fill = quartile)) +
  scale_x_continuous(breaks = quartiles, labels = round(quartiles, 0)) +
  # scale_x_continuous(n.breaks = 10) +
  scale_fill_viridis_d() +
  labs(title = "Observation density per patient",
       x = "Observation count", y = "Density", fill = "Quartile")

```

## Univariate plots

### Original dataset

#### Outcome

```{r univariate_plots_original_outcome}

na.omit(one_over_mac, cols = c("peak_alpha")) %>%
  ggplot(aes(x = peak_alpha)) +
  geom_density() +
  labs(x = "Peak alpha frequency",
       title = "Density of peak_alpha")

na.omit(one_over_mac, cols = c("peak_alpha")) %>%
  ggplot(aes(x = log(peak_alpha))) +
  geom_density() +
  labs(x = "Natural logarithm of peak alpha frequency",
       title = "Density of log(peak_alpha)")

na.omit(one_over_mac, cols = c("peak_alpha")) %>%
  ggplot(aes(x = 1 / peak_alpha)) +
  geom_density() +
  labs(x = "Inverse of peak alpha frequency",
       title = "Density of 1/peak_alpha")

```

#### Predictors

```{r univariate_plots_original_predictors}

na.omit(one_over_mac, cols = c("ce_mac")) %>%
  ggplot(aes(x = ce_mac)) +
  geom_density() +
  labs(x = "MAC at effect site", title = "Density of ce_mac")

na.omit(one_over_mac, cols = c("ce_mac")) %>%
  ggplot(aes(x = 1 / ce_mac)) +
  geom_density() +
  labs(x = "Inverse of MAC at effect site", title = "Density of 1/ce_mac") +
  scale_x_continuous(limits = c(0, 3))

na.omit(one_over_mac, cols = c("age")) %>%
  ggplot(aes(x = age)) +
  geom_density() +
  labs(x = "Age [years]", title = "Density of age")

```

### Filtered dataset

#### Outcome

```{r univariate_plots_filtered_outcome}

one_over_mac_filtered %>%
  ggplot(aes(x = peak_alpha)) +
  geom_density() +
  labs(x = "Peak alpha frequency", title = "Density of peak_alpha")

one_over_mac_filtered %>%
  ggplot(aes(x = log(peak_alpha))) +
  geom_density() +
  labs(x = "Natural logarithm of peak alpha frequency",
       title = "Density of log(peak_alpha)")

one_over_mac_filtered %>%
  ggplot(aes(x = 1/peak_alpha)) +
  geom_density() +
  labs(x = "Inverse of peak alpha frequency",
       title = "Density of 1/peak_alpha")
```

#### Predictors

```{r univariate_plots_filtered_predictors}

one_over_mac_filtered %>%
  ggplot(aes(x = ce_mac)) +
  geom_density() +
  labs(x = "MAC at effect site", title = "Density of ce_mac")

one_over_mac_filtered %>%
  ggplot(aes(x = 1/ce_mac)) +
  geom_density() +
  labs(x = "Inverse of MAC at effect site", title = "Density of 1/ce_mac") +
  scale_x_continuous(limits = c(0, 3))

one_over_mac_filtered %>%
  ggplot(aes(x = age)) +
  geom_density_line() +
  labs(x = "Age [years]", title = "Density of age")

```

## Bivariate plots

### Original dataset

#### Age versus ceMAC

```{r age_ce_mac_biv_original}

na.omit(one_over_mac, cols = c("age", "ce_mac")) %>%
  ggplot(aes(
    x = ce_mac,
    y = cut_width(age, width = 5),
    fill = factor(stat(quantile))
  )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantile_lines = TRUE,
    quantiles = 4,
    bandwidth = 0.05
  ) +
  labs(
    x = "MAC at effect site",
    y = "Age",
    title = "ceMAC by age",
    subtitle = "Fixed bin width [5]"
  ) + scale_fill_viridis_d(name = "Quartiles")

na.omit(one_over_mac, cols = c("age", "ce_mac")) %>%
  ggplot(aes(
    x = ce_mac,
    y = cut_number(age, n = 15),
    fill = factor(stat(quantile))
  )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantile_lines = TRUE,
    quantiles = 4,
    bandwidth = 0.05
  ) +
  labs(
    x = "MAC at effect site",
    y = "Age",
    title = "ceMAC by age",
    subtitle = "Equal number of observations per bin [15 bins]"
  ) + scale_fill_viridis_d(name = "Quartiles")

na.omit(one_over_mac, cols = c("age", "ce_mac")) %>%
  ggplot(aes(
    x = ce_mac,
    y = cut_interval(age, n = 15),
    fill = factor(stat(quantile))
  )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantile_lines = TRUE,
    quantiles = 4,
    bandwidth = 0.05
  ) +
  labs(
    x = "MAC at effect site",
    y = "Age",
    title = "ceMAC by age",
    subtitle = "Equal range [15 bins]"
  ) + scale_fill_viridis_d(name = "Quartiles")

```

#### Age versus peak alpha frequency

```{r age_peak_alpha_biv_original}

na.omit(one_over_mac, cols = c("age", "peak_alpha")) %>%
  ggplot(aes(
    x = peak_alpha,
    y = cut_width(age, width = 5),
    fill = factor(stat(quantile))
  )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantile_lines = TRUE,
    quantiles = 4,
    bandwidth = 0.25
  ) +
  labs(
    x = "Peak alpha frequency [Hz]",
    y = "Age",
    title = "Peak alpha frequency by age",
    subtitle = "Fixed bin width [5]"
  ) + scale_fill_viridis_d(name = "Quartiles")

na.omit(one_over_mac, cols = c("age", "peak_alpha")) %>%
  ggplot(aes(
    x = peak_alpha,
    y = cut_number(age, n = 15),
    fill = factor(stat(quantile))
  )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantile_lines = TRUE,
    quantiles = 4,
    bandwidth = 0.25
  ) +
  labs(
    x = "Peak alpha frequency [Hz]",
    y = "Age",
    title = "Peak alpha frequency by age",
    subtitle = "Equal number of observations per bin [15 bins]"
  ) + scale_fill_viridis_d(name = "Quartiles")

na.omit(one_over_mac, cols = c("age", "peak_alpha")) %>%
  ggplot(aes(
    x = peak_alpha,
    y = cut_interval(age, n = 15),
    fill = factor(stat(quantile))
  )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantile_lines = TRUE,
    quantiles = 4,
    bandwidth = 0.25
  ) +
  labs(
    x = "Peak alpha frequency [Hz]",
    y = "Age",
    title = "Peak alpha frequency by age",
    subtitle = "Equal range [15 bins]"
  ) + scale_fill_viridis_d(name = "Quartiles")

```

#### ceMAC versus peak alpha frequency

```{r ce_mac_peak_alpha_biv_original}

na.omit(one_over_mac, cols = c("ce_mac", "peak_alpha")) %>%
  ggplot(aes(
    x = peak_alpha,
    y = cut_width(ce_mac, width = 0.1),
    fill = factor(stat(quantile))
  )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantile_lines = TRUE,
    quantiles = 4
  ) +
  labs(
    x = "Peak alpha frequency [Hz]",
    y = "MAC at effect site",
    title = "Peak alpha frequency by ceMAC",
    subtitle = "Fixed bin width [0.1]"
  ) + scale_fill_viridis_d(name = "Quartiles")

na.omit(one_over_mac, cols = c("ce_mac", "peak_alpha")) %>%
  ggplot(aes(
    x = peak_alpha,
    y = cut_number(ce_mac, n = 25),
    fill = factor(stat(quantile))
  )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantile_lines = TRUE,
    quantiles = 4
  ) +
  labs(
    x = "Peak alpha frequency [Hz]",
    y = "MAC at effect site",
    title = "Peak alpha frequency by ceMAC",
    subtitle = "Equal number of observations per bin [25 bins]"
  ) + scale_fill_viridis_d(name = "Quartiles")

na.omit(one_over_mac, cols = c("ce_mac", "peak_alpha")) %>%
  ggplot(aes(
    x = peak_alpha,
    y = cut_interval(ce_mac, n = 25),
    fill = factor(stat(quantile))
  )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantile_lines = TRUE,
    quantiles = 4
  ) +
  labs(
    x = "Peak alpha frequency [Hz]",
    y = "MAC at effect site",
    title = "Peak alpha frequency by ceMAC",
    subtitle = "Equal range [25 bins]"
  ) + scale_fill_viridis_d(name = "Quartiles")

```

### Filtered dataset

#### Age versus ceMAC

```{r age_ce_mac_biv_filtered}

one_over_mac_filtered %>%
  ggplot(aes(
    x = ce_mac,
    y = cut_width(age, width = 5),
    fill = factor(stat(quantile))
  )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantile_lines = TRUE,
    quantiles = 4,
    bandwidth = 0.05
  ) +
  labs(
    x = "MAC at effect site",
    y = "Age",
    title = "ceMAC by age",
    subtitle = "Fixed bin width [5]"
  ) + scale_fill_viridis_d(name = "Quartiles")

one_over_mac_filtered %>%
  ggplot(aes(
    x = ce_mac,
    y = cut_number(age, n = 15),
    fill = factor(stat(quantile))
  )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantile_lines = TRUE,
    quantiles = 4,
    bandwidth = 0.05
  ) +
  labs(
    x = "MAC at effect site",
    y = "Age",
    title = "ceMAC by age",
    subtitle = "Equal number of observations per bin [15 bins]"
  ) + scale_fill_viridis_d(name = "Quartiles")

one_over_mac_filtered %>%
  ggplot(aes(
    x = ce_mac,
    y = cut_interval(age, n = 15),
    fill = factor(stat(quantile))
  )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantile_lines = TRUE,
    quantiles = 4,
    bandwidth = 0.05
  ) +
  labs(
    x = "MAC at effect site",
    y = "Age",
    title = "ceMAC by age",
    subtitle = "Equal range [25 bins]"
  ) + scale_fill_viridis_d(name = "Quartiles")

```

#### ceMAC versus peak alpha frequency

```{r ce_mac_peak_alpha_biv_filtered}

one_over_mac_filtered %>%
  ggplot(aes(
    x = peak_alpha,
    y = cut_width(ce_mac, width = 0.1),
    fill = factor(stat(quantile))
  )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantile_lines = TRUE,
    quantiles = 4
  ) +
  labs(
    x = "Peak alpha frequency [Hz]",
    y = "MAC at effect site",
    title = "Peak alpha frequency by ceMAC",
    subtitle = "Fixed bin width [0.1], filtered dataset"
  ) + scale_fill_viridis_d(name = "Quartiles")

one_over_mac_filtered %>%
  ggplot(aes(
    x = peak_alpha,
    y = cut_number(ce_mac, n = 10),
    fill = factor(stat(quantile))
  )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantile_lines = TRUE,
    quantiles = 4
  ) +
  labs(
    x = "Peak alpha frequency [Hz]",
    y = "MAC at effect site",
    title = "Peak alpha frequency by ceMAC",
    subtitle = "Equal number of observations per bin [10 bins], filtered dataset"
  ) + scale_fill_viridis_d(name = "Quartiles")

one_over_mac_filtered %>%
  ggplot(aes(
    x = peak_alpha,
    y = cut_interval(ce_mac, n = 10),
    fill = factor(stat(quantile))
  )) +
  stat_density_ridges(
    geom = "density_ridges_gradient",
    calc_ecdf = TRUE,
    quantile_lines = TRUE,
    quantiles = 4
  ) +
  labs(
    x = "Peak alpha frequency [Hz]",
    y = "MAC at effect site",
    title = "Peak alpha frequency by ceMAC",
    subtitle = "Equal range [10 bins], filtered dataset"
  ) + scale_fill_viridis_d(name = "Quartiles")

```

## Miscellaneous plots

```{r misc_plots}


```

## Autocorrelation plots

```{r autocorrelation}

acf <-
  one_over_mac_filtered[,
    .(data = list(.SD)),
    by = pid,
    .SDcols = c("ce_mac", "peak_alpha")
    ][,
    .(acf = map(data, ~ acf(
      .,
      lag.max = 2,
      type = "correlation",
      plot = FALSE
    )))
    ][,
    .(acf = map(
      acf,
      ~ data.table(
        ce_mac__lag_1 = .x$acf[2, 1, 1],
        peak_alpha__lag_1 = .x$acf[2, 2, 2],
        ce_mac__lag_2 = .x$acf[3, 1, 1],
        peak_alpha__lag_2 = .x$acf[3, 2, 2]
      )
    ))
    ][,
    lapply(.SD, rbindlist),
    .SDcols = c("acf")
    ]

acf <-
  melt(acf,
       measure.vars = patterns(ce_mac = "^acf.ce_mac__lag_",
                               peak_alpha = "^acf.peak_alpha__lag_"),
       variable.name = "lag")

acf <-
  melt(acf, id.vars = "lag", measure.vars = c("ce_mac", "peak_alpha"))

acf %>%
  ggplot(aes(x = as_factor(lag), y = value)) +
  facet_wrap(vars(variable),
             scales = "fixed",
             labeller = labeller(
               variable = c("ce_mac" = "ceMAC",
                            "peak_alpha" = "Peak alpha frequency")
             )) +
  geom_boxplot() +
  labs(x = "Lag", y = "Autocorrelation",
       title = "Autocorrelation by variable") +
  theme_dark()

```

## Partial autocorrelation plots

```{r partial_autocorrelation}

pacf <-
  one_over_mac_filtered[,
    .(data = list(.SD)),
    by = pid,
    .SDcols = c("ce_mac", "peak_alpha")
    ][,
    .(pacf = map(data, ~ pacf(
      .,
      lag.max = 2,
      plot = FALSE
    )))
    ][,
    .(pacf = map(
      pacf,
      ~ data.table(
      ce_mac__lag_1 = .x$acf[1, 1, 1],
      peak_alpha__lag_1 = .x$acf[1, 2, 2],
      ce_mac__lag_2 = .x$acf[2, 1, 1],
      peak_alpha__lag_2 = .x$acf[2, 2, 2]
    )
    ))
    ][,
    lapply(.SD, rbindlist),
    .SDcols = c("pacf")
    ]

pacf <-
  melt(pacf,
       measure.vars = patterns(ce_mac = "^pacf.ce_mac__lag_",
                               peak_alpha = "^pacf.peak_alpha__lag_"),
       variable.name = "lag")

pacf <-
  melt(pacf, id.vars = "lag", measure.vars = c("ce_mac", "peak_alpha"))

pacf %>%
  ggplot(aes(x = as_factor(lag), y = value)) +
  facet_wrap(vars(variable),
             scales = "fixed",
             labeller = labeller(
               variable = c("ce_mac" = "ceMAC",
                            "peak_alpha" = "Peak alpha frequency")
             )) +
  geom_boxplot() +
  labs(x = "Lag", y = "Partial autocorrelation",
       title = "Partial autocorrelation by variable") +
  theme_dark()

```
