---
title: "Modelling"
author: "Pascal Jerney"
date: "2021-03-08"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, echo=FALSE}

if (!require("pacman", quietly = TRUE)) {
  install.packages("pacman")
}
pacman::p_load("conflicted")

# Resolve function conflicts before loading other packages
conflict_prefer("select", "dplyr", quiet = TRUE)
conflict_prefer("filter", "dplyr", quiet = TRUE)
conflict_prefer("collapse", "dplyr", quiet = TRUE)

pacman::p_load(
  "tidyverse",
  "fs",
  "glue",
  "nlme",
  "MuMIn",
  "data.table"
)

options(scipen = 6, digits = 4) # I prefer to view outputs in non-scientific notation

```

## Introduction

Data is resampled groupwise over `pid` without replacing, with 100 observations
per patient and repeated 100 times. This is necessary to calculate models in a
timely manner.

## To do

* Nothing so far

## Modelling

### Null model

```{r load_null_model, include=FALSE}

model0_specs <-
  read_rds(file = path_wd("output", "model0_specs", ext = "rds"))

```

Intercept-only model with `log(peak_alpha)` as outcome.

#### Estimates plot

```{r null_model_plot}

model0_specs[, rbindlist(tidy)
  ][, .(estimate, mean = mean(estimate), median = median(estimate)), by = term
  ] %>%
  ggplot(aes(x = estimate)) +
  facet_wrap(vars(term), scales = "free") +
  geom_histogram(aes(y = ..density..), position = "identity", bins = 15) +
  geom_density() +
  geom_vline(aes(xintercept = mean),
             colour = "red",
             linetype = "dashed") +
  geom_vline(aes(xintercept = median),
             colour = "blue",
             linetype = "dashed") +
  labs(
    title = "Subsampled estimate plot",
    subtitle = "Mean in red, median in blue",
    x = "Estimate",
    y = "Density"
  )

```

#### Estimates table

```{r null_model_summary}

model0_estimates <-
  model0_specs[, rbindlist(tidy)
    ][, .(mean = mean(estimate),
          se = sd(estimate) / sqrt(.N)),
      by = term]

setkey(melt(model0_estimates, id.vars = "term"), term) %>%
  knitr::kable(digits = 10, col.names = c("Term", "Statistic", "Value"))

```

#### Performance

```{r null_model_performance}

model0_performance <-
  model0_specs[, rbindlist(performance, idcol = "seed")]

setkey(melt(model0_performance,
            id.vars = "seed",
            measure.vars = c("AIC", "RMSE")), seed, variable) %>%
  ggplot(aes(x = value)) +
  facet_wrap(
    vars(variable), scales = "free",
    labeller = labeller(variable = c("RMSE" = "RMSE", "AIC" = "AIC"))) +
  stat_density() +
  labs(
    title = "Subsampled performance plot",
    x = "Estimate",
    y = "Density"
  )

```

#### Residuals diagnostics

```{r null_model_resid_diag}

model0_residuals <-
  model0_specs[, rbindlist(residuals, idcol = "seed")]

model0_residuals[seed %in% c(1)] %>%
  ggplot(aes(sample = normalized)) +
  geom_qq() +
  geom_qq_line() +
  scale_colour_viridis_d() +
  labs(
    title = "Normalized residuals QQ-plot",
    # subtitle = "Colours indicate different subsample seeds",
    x = "Theoretical quantiles",
    y = "Sample quantiles"
  )

model0_residuals[seed %in% c(1:5)] %>%
  ggplot(aes(x = normalized, group = factor(seed), colour = factor(seed))) +
  geom_density() +
  scale_colour_viridis_d() +
  theme(legend.position = "none") +
  labs(
    title = "Subsampled normalized residuals density plot",
    subtitle = "Colours indicate different subsample seeds",
    x = "Normalized residuals",
    y = "Density"
  )

model0_acf <-
  model0_residuals[, .(residuals = list(.SD)), by = "seed"
    ][, .(seed, acf = map(residuals,
                   ~ acf(.x[, .(normalized)],
                         type = "correlation",
                         plot = FALSE)))]

model0_acf[1:5, rbindlist(acf, idcol = "seed")] %>%
  ggplot(aes(x = lag, y = acf, fill = factor(seed))) +
  geom_hline(aes(yintercept = 0)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_viridis_d() +
  theme(legend.position = "none") +
  labs(
    title = "Subsampled normalized residuals ACF plot",
    subtitle = "Colours indicate different subsample seeds",
    x = "Lag",
    y = "ACF"
  )

```

### Model with ce_mac as fixed effect

```{r load_linear_model, include=FALSE}

model1_specs <-
  read_rds(file = path_wd("output", "model1_specs", ext = "rds"))

```

Model with `ce_mac` as fixed effect and `log(peak_alpha)` as outcome.

#### Estimates plot

```{r linear_model_plot}

model1_specs[, rbindlist(tidy)
  ][, .(estimate, mean = mean(estimate), median = median(estimate)), by = term
  ] %>%
  ggplot(aes(x = estimate)) +
  facet_wrap(vars(term), scales = "free") +
  geom_histogram(aes(y = ..density..), position = "identity", bins = 15) +
  geom_density() +
  geom_vline(aes(xintercept = mean),
             colour = "red",
             linetype = "dashed") +
  geom_vline(aes(xintercept = median),
             colour = "blue",
             linetype = "dashed") +
  labs(
    title = "Subsampled estimate plot",
    subtitle = "Mean in red, median in blue",
    x = "Estimate",
    y = "Density"
  )

```

#### Estimates table

```{r linear_model_summary}

model1_estimates <-
  model1_specs[, rbindlist(tidy)
    ][, .(mean = mean(estimate),
          se = sd(estimate) / sqrt(.N)),
      by = term]

setkey(melt(model1_estimates, id.vars = "term"), term) %>%
  knitr::kable(digits = 10, col.names = c("Term", "Statistic", "Value"))

```

#### Performance

```{r linear_model_performance}

model1_performance <-
  model1_specs[, rbindlist(performance, idcol = "seed")]

setkey(melt(model1_performance,
            id.vars = "seed",
            measure.vars = c("AIC", "RMSE")), seed, variable) %>%
  ggplot(aes(x = value)) +
  facet_wrap(
    vars(variable), scales = "free",
    labeller = labeller(variable = c("RMSE" = "RMSE", "AIC" = "AIC"))) +
  stat_density() +
  labs(
    title = "Subsampled performance plot",
    x = "Estimate",
    y = "Density"
  )

```

#### Residuals diagnostics

```{r linear_model_resid_diag}

model1_residuals <-
  model1_specs[, rbindlist(residuals, idcol = "seed")]

model1_residuals[seed %in% c(1)] %>%
  ggplot(aes(sample = normalized)) +
  geom_qq() +
  geom_qq_line() +
  scale_colour_viridis_d() +
  labs(
    title = "Normalized residuals QQ-plot",
    # subtitle = "Colours indicate different subsample seeds",
    x = "Theoretical quantiles",
    y = "Sample quantiles"
  )

model1_residuals[seed %in% c(1:5)] %>%
  ggplot(aes(x = normalized, group = factor(seed), colour = factor(seed))) +
  geom_density() +
  scale_colour_viridis_d() +
  theme(legend.position = "none") +
  labs(
    title = "Subsampled normalized residuals density plot",
    subtitle = "Colours indicate different subsample seeds",
    x = "Normalized residuals",
    y = "Density"
  )

model1_acf <-
  model1_residuals[, .(residuals = list(.SD)), by = "seed"
    ][, .(seed, acf = map(residuals,
                   ~ acf(.x[, .(normalized)],
                         type = "correlation",
                         plot = FALSE)))]

model1_acf[1:5, rbindlist(acf, idcol = "seed")] %>%
  ggplot(aes(x = lag, y = acf, fill = factor(seed))) +
  geom_hline(aes(yintercept = 0)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_viridis_d() +
  theme(legend.position = "none") +
  labs(
    title = "Subsampled normalized residuals ACF plot",
    subtitle = "Colours indicate different subsample seeds",
    x = "Lag",
    y = "ACF"
  )

```

### Model with ce_mac as fixed effect and pid as random intercept

```{r load_mixed_intercept_model, include=FALSE}

model2_specs <-
  read_rds(file = path_wd("output", "model2_specs", ext = "rds"))

```

Model with `ce_mac` as fixed effect, `pid` as random intercept and
`log(peak_alpha)` as outcome.

#### Estimates plot

```{r mixed_intercept_model_plot}

model2_specs[, rbindlist(tidy)
  ][, .(estimate, mean = mean(estimate), median = median(estimate)), by = term
  ] %>%
  ggplot(aes(x = estimate)) +
  facet_wrap(vars(term), scales = "free") +
  geom_histogram(aes(y = ..density..), position = "identity", bins = 15) +
  geom_density() +
  geom_vline(aes(xintercept = mean),
             colour = "red",
             linetype = "dashed") +
  geom_vline(aes(xintercept = median),
             colour = "blue",
             linetype = "dashed") +
  labs(
    title = "Subsampled estimate plot",
    subtitle = "Mean in red, median in blue",
    x = "Estimate",
    y = "Density"
  )

```

#### Estimates table

```{r mixed_intercept_model_summary}

model2_estimates <-
  model2_specs[, rbindlist(tidy)
    ][, .(mean = mean(estimate),
          se = sd(estimate) / sqrt(.N)),
      by = term]

setkey(melt(model2_estimates, id.vars = "term"), term) %>%
  knitr::kable(digits = 10, col.names = c("Term", "Statistic", "Value"))

```

#### Performance

```{r mixed_intercept_model_performance}

model2_performance <-
  model2_specs[, rbindlist(performance, idcol = "seed")]

setkey(melt(model2_performance,
            id.vars = "seed",
            measure.vars = c("AIC", "RMSE")), seed, variable) %>%
  ggplot(aes(x = value)) +
  facet_wrap(
    vars(variable), scales = "free",
    labeller = labeller(variable = c("RMSE" = "RMSE", "AIC" = "AIC"))) +
  stat_density() +
  labs(
    title = "Subsampled performance plot",
    x = "Estimate",
    y = "Density"
  )

```

#### Residuals diagnostics

```{r mixed_intercept_model_resid_diag}

model2_residuals <-
  model2_specs[, rbindlist(residuals, idcol = "seed")]

model2_residuals[seed %in% c(1)] %>%
  ggplot(aes(sample = normalized)) +
  geom_qq() +
  geom_qq_line() +
  scale_colour_viridis_d() +
  labs(
    title = "Normalized residuals QQ-plot",
    # subtitle = "Colours indicate different subsample seeds",
    x = "Theoretical quantiles",
    y = "Sample quantiles"
  )

model2_residuals[seed %in% c(1:5)] %>%
  ggplot(aes(x = normalized, group = factor(seed), colour = factor(seed))) +
  geom_density() +
  scale_colour_viridis_d() +
  theme(legend.position = "none") +
  labs(
    title = "Subsampled normalized residuals density plot",
    subtitle = "Colours indicate different subsample seeds",
    x = "Normalized residuals",
    y = "Density"
  )

model2_acf <-
  model2_residuals[, .(residuals = list(.SD)), by = "seed"
    ][, .(seed, acf = map(residuals,
                   ~ acf(.x[, .(normalized)],
                         type = "correlation",
                         plot = FALSE)))]

model2_acf[1:5, rbindlist(acf, idcol = "seed")] %>%
  ggplot(aes(x = lag, y = acf, fill = factor(seed))) +
  geom_hline(aes(yintercept = 0)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_viridis_d() +
  theme(legend.position = "none") +
  labs(
    title = "Subsampled normalized residuals ACF plot",
    subtitle = "Colours indicate different subsample seeds",
    x = "Lag",
    y = "ACF"
  )

```

### Model with ce_mac as fixed effect and random intercept and slope of ce_mac on pid

```{r load_mixed_intercept_slope_model, include=FALSE}

model3_specs <-
  read_rds(file = path_wd("output", "model3_specs", ext = "rds"))

```

Model with `ce_mac` as fixed effect, `pid` as random intercept, `ce_mac:pid` as
random slope and `log(peak_alpha)` as outcome.

#### Estimates plot

```{r mixed_intercept_slope_model_plot}

model3_specs[, rbindlist(tidy)
  ][, .(estimate, mean = mean(estimate), median = median(estimate)), by = term
  ] %>%
  ggplot(aes(x = estimate)) +
  facet_wrap(vars(term), scales = "free") +
  geom_histogram(aes(y = ..density..), position = "identity", bins = 15) +
  geom_density() +
  geom_vline(aes(xintercept = mean),
             colour = "red",
             linetype = "dashed") +
  geom_vline(aes(xintercept = median),
             colour = "blue",
             linetype = "dashed") +
  labs(
    title = "Subsampled estimate plot",
    subtitle = "Mean in red, median in blue",
    x = "Estimate",
    y = "Density"
  )

```

#### Estimates table

```{r mixed_intercept_slope_model_summary}

model3_estimates <-
  model3_specs[, rbindlist(tidy)
    ][, .(mean = mean(estimate),
          se = sd(estimate) / sqrt(.N)),
      by = term]

setkey(melt(model3_estimates, id.vars = "term"), term) %>%
  knitr::kable(digits = 10, col.names = c("Term", "Statistic", "Value"))

```

#### Performance

```{r mixed_intercept_slope_model_performance}

model3_performance <-
  model3_specs[, rbindlist(performance, idcol = "seed")]

setkey(melt(model3_performance,
            id.vars = "seed",
            measure.vars = c("AIC", "RMSE")), seed, variable) %>%
  ggplot(aes(x = value)) +
  facet_wrap(
    vars(variable), scales = "free",
    labeller = labeller(variable = c("RMSE" = "RMSE", "AIC" = "AIC"))) +
  stat_density() +
  labs(
    title = "Subsampled performance plot",
    x = "Estimate",
    y = "Density"
  )

```

#### Residuals diagnostics

```{r mixed_intercept_slope_model_resid_diag}

model3_residuals <-
  model3_specs[, rbindlist(residuals, idcol = "seed")]

model3_residuals[seed %in% c(1)] %>%
  ggplot(aes(sample = normalized)) +
  geom_qq() +
  geom_qq_line() +
  scale_colour_viridis_d() +
  labs(
    title = "Normalized residuals QQ-plot",
    # subtitle = "Colours indicate different subsample seeds",
    x = "Theoretical quantiles",
    y = "Sample quantiles"
  )

model3_residuals[seed %in% c(1:5)] %>%
  ggplot(aes(x = normalized, group = factor(seed), colour = factor(seed))) +
  geom_density() +
  scale_colour_viridis_d() +
  theme(legend.position = "none") +
  labs(
    title = "Subsampled normalized residuals density plot",
    subtitle = "Colours indicate different subsample seeds",
    x = "Normalized residuals",
    y = "Density"
  )

model3_acf <-
  model3_residuals[, .(residuals = list(.SD)), by = "seed"
    ][, .(seed, acf = map(residuals,
                   ~ acf(.x[, .(normalized)],
                         type = "correlation",
                         plot = FALSE)))]

model3_acf[1:5, rbindlist(acf, idcol = "seed")] %>%
  ggplot(aes(x = lag, y = acf, fill = factor(seed))) +
  geom_hline(aes(yintercept = 0)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_viridis_d() +
  theme(legend.position = "none") +
  labs(
    title = "Subsampled normalized residuals ACF plot",
    subtitle = "Colours indicate different subsample seeds",
    x = "Lag",
    y = "ACF"
  )

```

### Model with ce_mac as fixed effect, random intercept and slope of ce_mac on pid and AR1 process within pid

```{r load_mixed_intercept_slope_ar1_model, include=FALSE}

model4_specs <-
  read_rds(file = path_wd("output", "model4_specs", ext = "rds"))

```

Model with `ce_mac` as fixed effect, `pid` as random intercept, `ce_mac:pid` as
random slope, AR1 error structure within `pid` and `log(peak_alpha)` as outcome.

#### Estimates plot

```{r mixed_intercept_slope_ar1_model_plot}

model4_specs[, rbindlist(tidy)
  ][, .(estimate, mean = mean(estimate), median = median(estimate)), by = term
  ] %>%
  ggplot(aes(x = estimate)) +
  facet_wrap(vars(term), scales = "free") +
  geom_histogram(aes(y = ..density..), position = "identity", bins = 15) +
  geom_density() +
  geom_vline(aes(xintercept = mean),
             colour = "red",
             linetype = "dashed") +
  geom_vline(aes(xintercept = median),
             colour = "blue",
             linetype = "dashed") +
  labs(
    title = "Subsampled estimate plot",
    subtitle = "Mean in red, median in blue",
    x = "Estimate",
    y = "Density"
  )

```

#### Estimates table

```{r mixed_intercept_slope_ar1_model_summary}

model4_estimates <-
  model4_specs[, rbindlist(tidy)
    ][, .(mean = mean(estimate),
          se = sd(estimate) / sqrt(.N)),
      by = term]

setkey(melt(model4_estimates, id.vars = "term"), term) %>%
  knitr::kable(digits = 10, col.names = c("Term", "Statistic", "Value"))

```

#### Performance

```{r mixed_intercept_slope_ar1_model_performance}

model4_performance <-
  model4_specs[, rbindlist(performance, idcol = "seed")]

setkey(melt(model4_performance,
            id.vars = "seed",
            measure.vars = c("AIC", "RMSE")), seed, variable) %>%
  ggplot(aes(x = value)) +
  facet_wrap(
    vars(variable), scales = "free",
    labeller = labeller(variable = c("RMSE" = "RMSE", "AIC" = "AIC"))) +
  stat_density() +
  labs(
    title = "Subsampled performance plot",
    x = "Estimate",
    y = "Density"
  )

```

#### Residuals diagnostics

```{r mixed_intercept_slope_ar1_model_resid_diag}

model4_residuals <-
  model4_specs[, rbindlist(residuals, idcol = "seed")]

model4_residuals[seed %in% c(1)] %>%
  ggplot(aes(sample = normalized)) +
  geom_qq() +
  geom_qq_line() +
  scale_colour_viridis_d() +
  labs(
    title = "Normalized residuals QQ-plot",
    # subtitle = "Colours indicate different subsample seeds",
    x = "Theoretical quantiles",
    y = "Sample quantiles"
  )

model4_residuals[seed %in% c(1:5)] %>%
  ggplot(aes(x = normalized, group = factor(seed), colour = factor(seed))) +
  geom_density() +
  scale_colour_viridis_d() +
  theme(legend.position = "none") +
  labs(
    title = "Subsampled normalized residuals density plot",
    subtitle = "Colours indicate different subsample seeds",
    x = "Normalized residuals",
    y = "Density"
  )

model4_acf <-
  model4_residuals[, .(residuals = list(.SD)), by = "seed"
    ][, .(seed, acf = map(residuals,
                   ~ acf(.x[, .(normalized)],
                         type = "correlation",
                         plot = FALSE)))]

model4_acf[1:5, rbindlist(acf, idcol = "seed")] %>%
  ggplot(aes(x = lag, y = acf, fill = factor(seed))) +
  geom_hline(aes(yintercept = 0)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_viridis_d() +
  theme(legend.position = "none") +
  labs(
    title = "Subsampled normalized residuals ACF plot",
    subtitle = "Colours indicate different subsample seeds",
    x = "Lag",
    y = "ACF"
  )

```

### Model with ce_mac as fixed effect, random intercept and slope of ce_mac on pid and continuous AR1 process within pid

```{r load_mixed_intercept_slope_car1_model, include=FALSE}

model5_specs <-
  read_rds(file = path_wd("output", "model5_specs", ext = "rds"))

```

Model with `ce_mac` as fixed effect, `pid` as random intercept, `ce_mac:pid` as
random slope, continuous AR1 error structure within `pid` and `log(peak_alpha)`
as outcome.

#### Estimates plot

```{r mixed_intercept_slope_car1_model_plot}

model5_specs[, rbindlist(tidy)
  ][, .(estimate, mean = mean(estimate), median = median(estimate)), by = term
  ] %>%
  ggplot(aes(x = estimate)) +
  facet_wrap(vars(term), scales = "free") +
  geom_histogram(aes(y = ..density..), position = "identity", bins = 15) +
  geom_density() +
  geom_vline(aes(xintercept = mean),
             colour = "red",
             linetype = "dashed") +
  geom_vline(aes(xintercept = median),
             colour = "blue",
             linetype = "dashed") +
  labs(
    title = "Subsampled estimate plot",
    subtitle = "Mean in red, median in blue",
    x = "Estimate",
    y = "Density"
  )

```

#### Estimates table

```{r mixed_intercept_slope_car1_model_summary}

model5_estimates <-
  model5_specs[, rbindlist(tidy)
    ][, .(mean = mean(estimate),
          se = sd(estimate) / sqrt(.N)),
      by = term]

setkey(melt(model5_estimates, id.vars = "term"), term) %>%
  knitr::kable(digits = 10, col.names = c("Term", "Statistic", "Value"))

```

#### Performance

```{r mixed_intercept_slope_car1_model_performance}

model5_performance <-
  model5_specs[, rbindlist(performance, idcol = "seed")]

setkey(melt(model5_performance,
            id.vars = "seed",
            measure.vars = c("AIC", "RMSE")), seed, variable) %>%
  ggplot(aes(x = value)) +
  facet_wrap(
    vars(variable), scales = "free",
    labeller = labeller(variable = c("RMSE" = "RMSE", "AIC" = "AIC"))) +
  stat_density() +
  labs(
    title = "Subsampled performance plot",
    x = "Estimate",
    y = "Density"
  )

```

#### Residuals diagnostics

```{r mixed_intercept_slope_car1_model_resid_diag}

model5_residuals <-
  model5_specs[, rbindlist(residuals, idcol = "seed")]

model5_residuals[seed %in% c(1)] %>%
  ggplot(aes(sample = normalized)) +
  geom_qq() +
  geom_qq_line() +
  scale_colour_viridis_d() +
  labs(
    title = "Normalized residuals QQ-plot",
    # subtitle = "Colours indicate different subsample seeds",
    x = "Theoretical quantiles",
    y = "Sample quantiles"
  )

model5_residuals[seed %in% c(1:5)] %>%
  ggplot(aes(x = normalized, group = factor(seed), colour = factor(seed))) +
  geom_density() +
  scale_colour_viridis_d() +
  theme(legend.position = "none") +
  labs(
    title = "Subsampled normalized residuals density plot",
    subtitle = "Colours indicate different subsample seeds",
    x = "Normalized residuals",
    y = "Density"
  )

model5_acf <-
  model5_residuals[, .(residuals = list(.SD)), by = "seed"
    ][, .(seed, acf = map(residuals,
                   ~ acf(.x[, .(normalized)],
                         type = "correlation",
                         plot = FALSE)))]

model5_acf[1:5, rbindlist(acf, idcol = "seed")] %>%
  ggplot(aes(x = lag, y = acf, fill = factor(seed))) +
  geom_hline(aes(yintercept = 0)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_viridis_d() +
  theme(legend.position = "none") +
  labs(
    title = "Subsampled normalized residuals ACF plot",
    subtitle = "Colours indicate different subsample seeds",
    x = "Lag",
    y = "ACF"
  )

```

### Model with ce_mac as fixed effect, random intercept and slope of ce_mac on pid and MA2 process within pid

```{r load_mixed_intercept_slope_ma2_model, include=FALSE}

model6_specs <-
  read_rds(file = path_wd("output", "model6_specs", ext = "rds"))

```

Model with `ce_mac` as fixed effect, `pid` as random intercept, `ce_mac:pid` as
random slope, MA2 error structure within `pid` and `log(peak_alpha)` as outcome.

#### Estimates plot

```{r mixed_intercept_slope_ma2_model_plot}

model6_specs[, rbindlist(tidy)
  ][, .(estimate, mean = mean(estimate), median = median(estimate)), by = term
  ] %>%
  ggplot(aes(x = estimate)) +
  facet_wrap(vars(term), scales = "free") +
  geom_histogram(aes(y = ..density..), position = "identity", bins = 15) +
  geom_density() +
  geom_vline(aes(xintercept = mean),
             colour = "red",
             linetype = "dashed") +
  geom_vline(aes(xintercept = median),
             colour = "blue",
             linetype = "dashed") +
  labs(
    title = "Subsampled estimate plot",
    subtitle = "Mean in red, median in blue",
    x = "Estimate",
    y = "Density"
  )

```

#### Estimates table

```{r mixed_intercept_slope_ma2_model_summary}

model6_estimates <-
  model6_specs[, rbindlist(tidy)
    ][, .(mean = mean(estimate),
          se = sd(estimate) / sqrt(.N)),
      by = term]

setkey(melt(model6_estimates, id.vars = "term"), term) %>%
  knitr::kable(digits = 10, col.names = c("Term", "Statistic", "Value"))

```

#### Performance

```{r mixed_intercept_slope_ma2_model_performance}

model6_performance <-
  model6_specs[, rbindlist(performance, idcol = "seed")]

setkey(melt(model6_performance,
            id.vars = "seed",
            measure.vars = c("AIC", "RMSE")), seed, variable) %>%
  ggplot(aes(x = value)) +
  facet_wrap(
    vars(variable), scales = "free",
    labeller = labeller(variable = c("RMSE" = "RMSE", "AIC" = "AIC"))) +
  stat_density() +
  labs(
    title = "Subsampled performance plot",
    x = "Estimate",
    y = "Density"
  )

```

#### Residuals diagnostics

```{r mixed_intercept_slope_ma2_model_resid_diag}

model6_residuals <-
  model6_specs[, rbindlist(residuals, idcol = "seed")]

model6_residuals[seed %in% c(1)] %>%
  ggplot(aes(sample = normalized)) +
  geom_qq() +
  geom_qq_line() +
  scale_colour_viridis_d() +
  labs(
    title = "Normalized residuals QQ-plot",
    # subtitle = "Colours indicate different subsample seeds",
    x = "Theoretical quantiles",
    y = "Sample quantiles"
  )

model6_residuals[seed %in% c(1:5)] %>%
  ggplot(aes(x = normalized, group = factor(seed), colour = factor(seed))) +
  geom_density() +
  scale_colour_viridis_d() +
  theme(legend.position = "none") +
  labs(
    title = "Subsampled normalized residuals density plot",
    subtitle = "Colours indicate different subsample seeds",
    x = "Normalized residuals",
    y = "Density"
  )

model6_acf <-
  model6_residuals[, .(residuals = list(.SD)), by = "seed"
    ][, .(seed, acf = map(residuals,
                   ~ acf(.x[, .(normalized)],
                         type = "correlation",
                         plot = FALSE)))]

model6_acf[1:5, rbindlist(acf, idcol = "seed")] %>%
  ggplot(aes(x = lag, y = acf, fill = factor(seed))) +
  geom_hline(aes(yintercept = 0)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_viridis_d() +
  theme(legend.position = "none") +
  labs(
    title = "Subsampled normalized residuals ACF plot",
    subtitle = "Colours indicate different subsample seeds",
    x = "Lag",
    y = "ACF"
  )

```

### Model with ce_mac as fixed effect, random intercept and slope of ce_mac on pid and ARMA1,1 process within pid

```{r load_mixed_intercept_slope_arma11_model, include=FALSE}

model7_specs <-
  read_rds(file = path_wd("output", "model7_specs", ext = "rds"))

```

Model with `ce_mac` as fixed effect, `pid` as random intercept, `ce_mac:pid` as
random slope, ARMA1,1 error structure within `pid` and `log(peak_alpha)` as
outcome.

#### Estimates plot

```{r mixed_intercept_slope_arma11_model_plot}

model7_specs[, rbindlist(tidy)
  ][, .(estimate, mean = mean(estimate), median = median(estimate)), by = term
  ] %>%
  ggplot(aes(x = estimate)) +
  facet_wrap(vars(term), scales = "free") +
  geom_histogram(aes(y = ..density..), position = "identity", bins = 15) +
  geom_density() +
  geom_vline(aes(xintercept = mean),
             colour = "red",
             linetype = "dashed") +
  geom_vline(aes(xintercept = median),
             colour = "blue",
             linetype = "dashed") +
  labs(
    title = "Subsampled estimate plot",
    subtitle = "Mean in red, median in blue",
    x = "Estimate",
    y = "Density"
  )

```

#### Estimates table

```{r mixed_intercept_slope_arma11_model_summary}

model7_estimates <-
  model7_specs[, rbindlist(tidy)
    ][, .(mean = mean(estimate),
          se = sd(estimate) / sqrt(.N)),
      by = term]

setkey(melt(model7_estimates, id.vars = "term"), term) %>%
  knitr::kable(digits = 10, col.names = c("Term", "Statistic", "Value"))

```

#### Performance

```{r mixed_intercept_slope_arma11_model_performance}

model7_performance <-
  model7_specs[, rbindlist(performance, idcol = "seed")]

setkey(melt(model7_performance,
            id.vars = "seed",
            measure.vars = c("AIC", "RMSE")), seed, variable) %>%
  ggplot(aes(x = value)) +
  facet_wrap(
    vars(variable), scales = "free",
    labeller = labeller(variable = c("RMSE" = "RMSE", "AIC" = "AIC"))) +
  stat_density() +
  labs(
    title = "Subsampled performance plot",
    x = "Estimate",
    y = "Density"
  )

```

#### Residuals diagnostics

```{r mixed_intercept_slope_arma11_model_resid_diag}

model7_residuals <-
  model7_specs[, rbindlist(residuals, idcol = "seed")]

model7_residuals[seed %in% c(1)] %>%
  ggplot(aes(sample = normalized)) +
  geom_qq() +
  geom_qq_line() +
  scale_colour_viridis_d() +
  labs(
    title = "Normalized residuals QQ-plot",
    # subtitle = "Colours indicate different subsample seeds",
    x = "Theoretical quantiles",
    y = "Sample quantiles"
  )

model7_residuals[seed %in% c(1:5)] %>%
  ggplot(aes(x = normalized, group = factor(seed), colour = factor(seed))) +
  geom_density() +
  scale_colour_viridis_d() +
  theme(legend.position = "none") +
  labs(
    title = "Subsampled normalized residuals density plot",
    subtitle = "Colours indicate different subsample seeds",
    x = "Normalized residuals",
    y = "Density"
  )

model7_acf <-
  model7_residuals[, .(residuals = list(.SD)), by = "seed"
    ][, .(seed, acf = map(residuals,
                   ~ acf(.x[, .(normalized)],
                         type = "correlation",
                         plot = FALSE)))]

model7_acf[1:5, rbindlist(acf, idcol = "seed")] %>%
  ggplot(aes(x = lag, y = acf, fill = factor(seed))) +
  geom_hline(aes(yintercept = 0)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_viridis_d() +
  theme(legend.position = "none") +
  labs(
    title = "Subsampled normalized residuals ACF plot",
    subtitle = "Colours indicate different subsample seeds",
    x = "Lag",
    y = "ACF"
  )

```

## Compare model performances

```{r performance_comparison}

# Compare performances plot
setkey(melt(rbind(
  model0_specs,
  model1_specs,
  model2_specs,
  model3_specs,
  model4_specs,
  model5_specs,
  model6_specs,
  model7_specs,
  idcol = "model")[, .(seed,
                       model,
                       rbindlist(performance, fill = TRUE))],
  id.vars = c("seed", "model"),
  measure.vars = c("AIC", "RMSE")),
  model, seed, variable)[] %>%
  ggplot(aes(x = value,
             fill = factor(model,
                           labels = c("Null",
                                      "Linear",
                                      "Mixed intercept",
                                      "Mixed intercept/slope",
                                      "Mixed intercept/slope AR1",
                                      "Mixed intercept/slope CAR1",
                                      "Mixed intercept/slope MA2",
                                      "Mixed intercept/slope ARMA1,1")))) +
  facet_wrap(vars(variable), scales = "free", ncol = 1,
  labeller = labeller(variable = c("RMSE" = "RMSE", "AIC" = "AIC"))) +
  stat_density(alpha = 0.5, position = position_identity()) +
  scale_fill_viridis_d() +
  labs(
    title = "Subsampled performance density plot",
    fill = "Model",
    x = "Value",
    y = "Density"
  )

```
